---
title: "Filtering mobile phone spam with the Naive Bayes algorithm"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: github_document
---


```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Information
* Example – filtering mobile phone spam with the Naive Bayes algorithm
* Data downloaded from http://www.dt.fee.unicamp.br/~tiago/smsspamcollection/smsspamcollection.zip.
* 将数据另存为sms_spam.csv


## Step 1. Prepare data

* 读取数据
```{r, eval=T, echo = T, tidy = TRUE, size="small", cache=FALSE, warning=FALSE}
sms_raw <- read.csv("smsspamcollection/sms_spam.csv", stringsAsFactors = FALSE)
sms_raw$type <- factor(sms_raw$type)

str(sms_raw)
```

* Cleaning and standardizing text data
```{r, eval=T, echo = T, tidy = TRUE, size="small", cache=FALSE, warning=FALSE}
library(SnowballC)
library(tm)

sms_corpus <- VCorpus(VectorSource(sms_raw$text))

sms_corpus_clean <- tm_map(sms_corpus, content_transformer(tolower))
sms_corpus_clean <- tm_map(sms_corpus,removeNumbers)
sms_corpus_clean <- tm_map(sms_corpus_clean, removeWords, stopwords())
sms_corpus_clean <- tm_map(sms_corpus_clean, removePunctuation)

sms_corpus_clean <- tm_map(sms_corpus_clean, stemDocument)
sms_corpus_clean <- tm_map(sms_corpus_clean, stripWhitespace)

sms_corpus_clean2 <- lapply(sms_corpus_clean[1:5572], as.character)
as.character(sms_corpus_clean2[1:3])
```

* Splitting text documents into words
```{r, eval=T, echo = T, tidy = TRUE, size="small", cache=FALSE, warning=FALSE}
sms_dtm <- DocumentTermMatrix(sms_corpus_clean)
sms_dtm

```

* Creating training and test datasets
```{r, eval=T, echo = T, tidy = TRUE, size="small", cache=FALSE, warning=FALSE}
sms_dtm_train <- sms_dtm[1:4169, ] 
sms_dtm_test <- sms_dtm[4170:5559, ]

sms_train_labels <- sms_raw[1:4169, ]$type 
sms_test_labels <- sms_raw[4170:5559, ]$type

prop.table(table(sms_train_labels))
prop.table(table(sms_test_labels))

```

* Visualizing text data – word clouds
```{r, eval=T, echo = T, tidy = TRUE, size="small", cache=FALSE, warning=FALSE}
library(wordcloud)
wordcloud(sms_corpus_clean, min.freq = 50, random.order = F, colors = rainbow(10))
```

* Visualizing spam word clouds
```{r, eval=T, echo = T, tidy = TRUE, size="small", cache=FALSE, warning=FALSE}
#ham <- subset(sms_raw, type == "ham")
#wordcloud(ham$text, max.words = 40, scale = c(3, 0.5), colors = rainbow(10))

spam <- subset(sms_raw, type == "spam")
wordcloud(spam$text, max.words = 40, scale = c(3, 0.5), colors = rainbow(10))
```

## Step 2. Training a model on the data

* Creating indicator features for frequent words
```{r, eval=T, echo = T, tidy = TRUE, size="small", cache=FALSE, warning=FALSE}
sms_freq_words <- findFreqTerms(sms_dtm_train, 5)
sms_dtm_freq_train<- sms_dtm_train[ , sms_freq_words]
sms_dtm_freq_test <- sms_dtm_test[ , sms_freq_words]

convert_counts <- function(x) { 
  x <- ifelse(x > 0, "Yes", "No")
}

sms_train <- apply(sms_dtm_freq_train, MARGIN = 2, convert_counts)
sms_test <- apply(sms_dtm_freq_test, MARGIN = 2, convert_counts)

library(e1071)
sms_classifier <- naiveBayes(sms_train, sms_train_labels)

```

## Step 3. Evaluating model performance

```{r, eval=T, echo = T, tidy = TRUE, size="small", cache=FALSE, warning=FALSE}
sms_test_pred <- predict(sms_classifier, sms_test)
library(gmodels)
CrossTable(sms_test_pred, sms_test_labels, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))

```

## Step 4. Improving model performance

```{r, eval=T, echo = T, tidy = TRUE, size="small", cache=FALSE, warning=FALSE}
sms_classifier2 <- naiveBayes(sms_train, sms_train_labels, laplace = 1)
sms_test_pred2 <- predict(sms_classifier2, sms_test)

CrossTable(sms_test_pred2, sms_test_labels, prop.chisq = FALSE, prop.t = FALSE, prop.r = FALSE, dnn = c('predicted', 'actual'))

```