---
title: "Diagnosing breast cancer with the k-NN algorithm"
output: github_document
---


```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Information
Diagnosing breast cancer with the **k-NN** algorithm, data downloaded from  http://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer-wisconsin/.


## Step 1. Download  and prepare data

* 直接从网站获取数据
```{r, eval=T, echo = T, tidy = TRUE, size="small", cache=FALSE, warning=FALSE}
wbcd <- read.table("http://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer-wisconsin/wdbc.data", sep = ",", header = F, stringsAsFactors = FALSE)
colnames(wbcd)<-c("id","diagnosis","radius_mean","texture_mean","perimeter_mean","area_mean")
head(wbcd)
```

## Step 2. Transformation – normalizing numeric data

* 癌症数据均一化处理
```{r, eval=T, echo = T, tidy = TRUE, size="small", cache=FALSE, warning=FALSE}
wbcd <- wbcd[-1]
wbcd$diagnosis<- factor(wbcd$diagnosis, levels = c("B", "M"), labels = c("Benign", "Malignant"))
#round(prop.table(table(wbcd$diagnosis)) * 100, digits = 1)
#summary(wbcd[c("radius_mean", "area_mean")])

normalize <- function(x){
  return ((x - min(x)) / (max(x) - min(x)))
}

wbcd_n <- as.data.frame(lapply(wbcd[2:31], normalize))
summary(wbcd_n$area_mean)
```

## step 3. Training a model on the data

* 训练模型
```{r, eval=T, echo = T, tidy = TRUE, size="small", cache=FALSE, warning=FALSE}
wbcd_train <- wbcd_n[1:469, ]
wbcd_test <- wbcd_n[470:569, ]

wbcd_train_labels <- wbcd[1:469, 1] 
wbcd_test_labels <- wbcd[470:569, 1]

library(class)
wbcd_test_pred <- knn(train = wbcd_train, test = wbcd_test, cl = wbcd_train_labels, k = 21)

```

## step 4. Evaluating model performance 

* 评估数据模型
```{r, eval=T, echo = T, tidy = TRUE, size="small", cache=FALSE, warning=FALSE}
library(gmodels)
CrossTable(x = wbcd_test_labels, y = wbcd_test_pred, prop.chisq=FALSE)
```
错误率仅为2%，即该模型可准确鉴定98%的肿瘤；

## step 5. Improving model performance

* 提升模型鉴别能力
```{r, eval=T, echo = T, tidy = TRUE, size="small", cache=FALSE, warning=FALSE}
# z-score standardization
wbcd_z <- as.data.frame(scale(wbcd[-1]))
summary(wbcd_z$area_mean)    # To confirm that the transformation was applied correctly;

wbcd_train <- wbcd_z[1:469, ] 
wbcd_test <- wbcd_z[470:569, ] 
wbcd_train_labels <- wbcd[1:469, 1] 
wbcd_test_labels <- wbcd[470:569, 1]

wbcd_test_pred <- knn(train = wbcd_train, test = wbcd_test, cl = wbcd_train_labels, k = 21)
CrossTable(x = wbcd_test_labels, y = wbcd_test_pred, prop.chisq = FALSE)

```
z-score标准化和normalize均一化数据无差异；

## step 6. Testing alternative values of k

* 尝试多个k值评估错误率

### k = 1
```{r, eval=T, echo = T, tidy = TRUE, size="small", cache=FALSE, warning=FALSE}

wbcd_test_pred_1 <- knn(train = wbcd_train, test = wbcd_test, cl = wbcd_train_labels, k = 1)
CrossTable(x = wbcd_test_labels, y = wbcd_test_pred_1, prop.chisq = FALSE)
```

### k = 5
```{r, eval=T, echo = T, tidy = TRUE, size="small", cache=FALSE, warning=FALSE}

wbcd_test_pred_5 <- knn(train = wbcd_train, test = wbcd_test, cl = wbcd_train_labels, k = 5)
CrossTable(x = wbcd_test_labels, y = wbcd_test_pred_5, prop.chisq = FALSE)
```

### k = 11
```{r, eval=T, echo = T, tidy = TRUE, size="small", cache=FALSE, warning=FALSE}

wbcd_test_pred_11 <- knn(train = wbcd_train, test = wbcd_test, cl = wbcd_train_labels, k = 11)
CrossTable(x = wbcd_test_labels, y = wbcd_test_pred_11, prop.chisq = FALSE)
```

### k = 15
```{r, eval=T, echo = T, tidy = TRUE, size="small", cache=FALSE, warning=FALSE}

wbcd_test_pred_15 <- knn(train = wbcd_train, test = wbcd_test, cl = wbcd_train_labels, k = 15)
CrossTable(x = wbcd_test_labels, y = wbcd_test_pred_15, prop.chisq = FALSE)
```

### k = 27
```{r, eval=T, echo = T, tidy = TRUE, size="small", cache=FALSE, warning=FALSE}

wbcd_test_pred_27 <- knn(train = wbcd_train, test = wbcd_test, cl = wbcd_train_labels, k = 27)
CrossTable(x = wbcd_test_labels, y = wbcd_test_pred_27, prop.chisq = FALSE)
```


